A Phenomenological Model for the Intracluster Medium that matches X-ray and Sunyaev-Zel'dovich observations

### 2013-10-02 paper1.tex
small clean-ups

{\bf We build a complete cosmological sample of galaxy clusters from the
  MultiDark $N$-body simulation with redshifts ranging from $z = 0$ to 1 and
  construct a \emph{phenomenological} model for the ICM. This is characterized
  by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample)
  and a cluster mass-dependent gas fraction. Note that our model is calibrated
  off X-ray observables within $R_{500}$. More work would be needed to
  extrapolate the model to larger cluster radii that are characterised by an
  increased complexity, which manifests itself in a larger kinetic-to-thermal
  pressure support, an increased level of density clumping, as well as cluster
  asphericity and ellipticity. In our model, we assign a (cluster
  mass-dependent) gas density profile to each DM halo in our sample and sort it
  into the NCCC/CCC populations according to a dynamical disturbance parameter
  that is calculated from the DM distribution. Applying the model to our sample
  of cluster halos, we obtain volume complete mock catalogs for the thermal (SZ
  and X-ray) cluster emission. Our mock catalogs match the observed X-ray
  luminosity function as well as the $L_{\rmn{X, bol}}-M_{500}$,
  $Y_{\rmn{X}}-M_{500}$, and $Y_{\rmn{SZ}}-M_{500}$ relations well.

  However, there are some notable deviations among the different observational
  scaling relations and our model, which may either point to observational
  sample selection effects that are not accounted for or missing complexity of
  our model. This model was specifically constructed to provide a reliable
  description of the gas density, which necessarily implies a great match of the
  mock X-ray observables (scaling relations and luminosity functions) as well as
  the low-redshift and massive cluster sample probed by {\em Planck}. However,
  we note that our model has a somewhat shallower slope of the
  $Y_{\rmn{SZ}}-M_{500}$ relation, which implies that the X-ray inferred
  deviations of self-similarity on small scales become weaker on the larger
  scales probed through the SZ effect. Moreover, our model overpredicts the
  normalization of the ACT $Y_{\rmn{SZ}}-M_{500}$ relation (at $z=0.53$) by
  about 30 per cent. This could either indicate possible selection effects of
  the different cluster samples or may even signal deviations from the assumed
  self-similar redshift evolution.

  In the companion Paper~II, we additionally present a model for the radio and
  gamma-ray emission as a result of hadronic CR interactions with the ICM.  We
  provide cosmologically complete multi-frequency mock catalogs for the
  \mbox{(non-)thermal} cluster emission at different redshifts and make these
  catalogs publicly and freely available on-line through the MultiDark database
  (www.multidark.org). Our mock catalogs are based on density profile shapes
  that represent the average cluster population of cool-core and non cool-core
  systems (while we account for scatter in the gas mass fractions, i.e., the
  normalization of the profiles). As a result, our model does not fully exploit
  the true variance across the average profile shapes which may result in a
  limited variance of thermal cluster observables.  Hence, the mock catalogs
  will allow quantitative comparisons to future hydrodynamical simulations and
  forecasts for future surveys.  We emphasize, however, that the catalogs are
  not suited for detailed cosmological studies in their (present) published
  form, which do not address any survey response functions nor provide
  light-cone mock catalogs. However, future surveys should be able to test the
  underlying assumptions of our modeling approach, which is based on X-ray
  observations of clusters at low redshift ($z<0.2$) and adopts self-similar
  redshift evolution for extrapolating to higher redshifts. We hope that the
  community can make valuable use of these catalogs in synergy with the future
  radio, X-ray and gamma-ray data.}

>>>>

{\bf We build a complete cosmological sample of galaxy clusters from the
  MultiDark $N$-body simulation with redshifts ranging from $z = 0$ to 1 and
  construct a \emph{phenomenological} model for the ICM. This is characterized
  by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample)
  and a cluster mass-dependent gas fraction. Note that our model is calibrated
  off X-ray observables within $R_{500}$. More work would be needed to
  extrapolate the model to larger cluster radii that are characterised by an
  increased complexity, which manifests itself in a larger kinetic-to-thermal
  pressure support, an increased level of density clumping, as well as cluster
  asphericity and ellipticity. In our model, we assign a (cluster
  mass-dependent) gas density profile to each DM halo in our sample and sort it
  into NCCC/CCC populations according to a dynamical disturbance parameter
  that is calculated from the DM distribution. Applying the model to our sample
  of cluster halos, we obtain volume-complete mock catalogs for the thermal (SZ
  and X-ray) cluster emission. Our mock catalogs match the observed X-ray
  luminosity function as well as the $L_{\rmn{X, bol}}-M_{500}$,
  $Y_{\rmn{X}}-M_{500}$, and $Y_{\rmn{SZ}}-M_{500}$ relations well.

  However, there are some notable deviations among the different observational
  scaling relations and our model, which may either point to observational
  sample selection effects that are not accounted for or missing complexity of
  our model. This model was specifically constructed to provide a reliable
  description of the gas density, which necessarily implies a great match of the
  resulting mock X-ray observables to data (scaling relations and luminosity
  functions) as well as to the low-redshift and massive cluster sample probed by
  {\em Planck}. However, we note that our model has a somewhat shallower slope
  of the $Y_{\rmn{SZ}}-M_{500}$ relation, which implies that the X-ray inferred
  deviations of self-similarity on small scales become weaker on the larger
  scales probed through the SZ effect. Moreover, our model overpredicts the
  normalization of the ACT $Y_{\rmn{SZ}}-M_{500}$ relation (at $z=0.53$) by
  about 30 per cent. This could either indicate possible selection effects of
  the different cluster samples or may even signal a deviation from the assumed
  self-similar redshift evolution.

  In the companion Paper~II, we additionally present a model for the radio and
  gamma-ray emission as a result of hadronic CR interactions with the ICM.  We
  provide cosmologically complete multi-frequency mock catalogs for the
  \mbox{(non-)thermal} cluster emission at different redshifts and make these
  catalogs publicly and freely available on-line through the MultiDark database
  (www.multidark.org). Our mock catalogs are based on density profile shapes
  that represent the {\em average} cluster population of cool-core and non
  cool-core systems (while we account for scatter in the gas mass fractions,
  i.e., the normalization of the profiles). As a result, our model does not
  fully exploit the true variance across the average profile shapes which may
  result in a limited variance of thermal cluster observables. Nevertheless, the
  mock catalogs will allow quantitative comparisons to future hydrodynamical
  simulations and forecasts for future surveys.  We emphasize, however, that the
  catalogs are not suited for detailed cosmological studies in their (present)
  published form, which do not address any survey response functions nor provide
  light-cone mock catalogs. However, future surveys should be able to test the
  underlying assumptions of our modeling approach, which is based on X-ray
  observations of clusters at low redshift ($z<0.2$) and adopts self-similar
  redshift evolution for extrapolating to higher redshifts. We hope that the
  community can make valuable use of these catalogs in synergy with the future
  radio, X-ray and gamma-ray data.}
